Modular Model-Checking of a Byzantine Fault-Tolerant Protocol

### 2016-12-08 doc/om-paper/paper.tex
final-ish proof read for sections 4+

The use of abstract transition systems not only allows us to scale proofs farther, but also to improve traceability and debugging while developing a model. In models like the ones described in Section~\ref{sec:modular} where there are on the order of 100 state variables and counter-example traces could be 30 steps long, the designer can be easily lost trying to identify the essence. In such cases, the values of the abstract predicates can serve to focus the designer's attention on one particular mode of the system where the counter example is taking place.

@@

% Intuitively, there are subsets of the overall state space which are important to understanding the evolution of such a system. For example, let $A_1$ denote the subset of states in which the processor $S$ has not sent a message, let $A_2$ be the subset in which $S$ has broadcast a message, and let $A_3$ be the subset in which $L_1$ and $L_2$ have each received and processed the message. The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counter-example traces.

@@

To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs. The results are summarized in Figure~\ref{fig:effort} and sketched below. In the table, for each modification, we report how much of the model must be modified. We report on four aspects of the system: which transition systems are modified (as described in Section~\ref{sec:sketch}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to. We modify the implementation along the axes of faults, time, and local node behavior.

@@

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The ADSL should be able to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that the ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}

>>>>

The use of abstract transition systems not only allows us to scale proofs farther, but also to improve traceability and debugging while developing a model. In models like the ones described in Section~\ref{sec:modular} where there are on the order of 100 state variables and counterexample traces could be 30 steps long, the designer can be easily lost trying to identify the essence. In such cases, the values of the abstract predicates can serve to focus the designer's attention on one particular mode of the system where the counter example is taking place.

@@

% Intuitively, there are subsets of the overall state space which are important to understanding the evolution of such a system. For example, let $A_1$ denote the subset of states in which the processor $S$ has not sent a message, let $A_2$ be the subset in which $S$ has broadcast a message, and let $A_3$ be the subset in which $L_1$ and $L_2$ have each received and processed the message. The union $A_1 \cup A_2 \cup A_3$ covers the set of reachable states of the system and defines an ASM that is useful for reasoning about its dynamics and understanding counterexample traces.

@@

To demonstrate the modularity of the modelling and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs. The results are summarized in Figure~\ref{fig:effort} and sketched below. In the table, for each modification, we report how much of the model must be modified. We report on four aspects of the system: which transition systems are modified (as described in Section~\ref{sec:sketch}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to. We modify the implementation along the axes of faults, time, and local node behavior.

@@

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The ADSL should be able to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that the ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models in SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
